---
title: "Creating_custome_settleR_plots"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Creating_custome_settleR_plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
# Introduction 
## Background and Motivation 
I had a need to convey combinations of sets and found the the usual method to do this, venn-diagrams or Euler plots too limiting. My research invovles identifying differentially expressed genes which involved looking large numbers of datasets which could be grouped any number of ways. I found [UpSet plots](https://github.com/hms-dbmi/UpSetR/) to offer a succint way to visualize these comparisons that is relatively intuitive to my audience. There is already an [UpSet package](https://cran.rstudio.com/web/packages/UpSetR/), however, I found it was not amenable to customization. In my experience, it is very difficult to write re-usable, modular code creating plots for anything besides one-off functions that will be used once, or at most a single context. 

## settleR: A simple and customizable package for creating UpSet plots 
I created the settleR package to to address the percieved challenges primarily for my own needs, but I hope that others may find it useful too. 

```{r}

```




```{r setup}
library(settleR)
```
