---
title: "Introduction_to_settleR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction_to_settleR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction 
# Background and Motivation 
I had a need to convey combinations of sets and found the the usual method to do this, venn-diagrams or Euler plots too limiting. My research invovles identifying differentially expressed genes which involved looking large numbers of datasets which could be grouped any number of ways. I found [UpSet plots](https://github.com/hms-dbmi/UpSetR/) to offer a succint way to visualize these comparisons which is relatively intuitive to my audience. There is already an [UpSet package](https://cran.rstudio.com/web/packages/UpSetR/), however, I found it was not amenable to customization. In my experience, it is very difficult to write re-usable, modular code creating plots for anything besides one-off functions that will be used once, or at most a single context. 

# settleR: A simple and customizable package for creating UpSet plots 
I created the settleR package to to address the percieved challenges primarily for my own needs, but I hope that others may find it useful too. 
The goal of settleR is to:

1. Create a simple/modular framework for UpSet-like figures
2. Make creating figures as painless as possible 
3. Allow for arbitrary customization 


```{r setup}
library(settleR)
library(grid)

```
# Creating a basic plot 
```{r, out.width="50%"}
# Loads an example setlist 
gene_setlist <- 
  system.file('extdata','ex_gene_setlist.rds', package = 'settleR') %>% 
  readRDS(.)

p <-  settleR_plot(gene_setlist)
grid.draw(p)
```


## Adding color to the groups 
```{r}
col_map <- 
  system.file('extdata','ex_col_map.rds', package = 'settleR') %>% 
  readRDS(.)

```
## Adding boxes around specific groups 
```{r}
setLevels <- 
  system.file('extdata','ex_set_levels.rds', package = 'settleR') %>% 
  readRDS(.)

intersect_lvls <- get_reordered_intersect_lvls(gene_setlist,
                                               setLevels = setLevels,
                                               nintersects = 15)

bm <- sets_to_matrix(gene_setlist) %>% 
  calc_set_overlaps(., setLevels = setLevels, 
                    intersectLevels = intersect_lvls) 
  
ul <- make_upset_plots(bm, col_map = col_map)
bound_boxes <- intersect_lvls[c(10:13)] %>% 
  box_intercepts_dims(ul$pmain, .)
pmain <- ul$pmain + 
  geom_rect(data=bound_boxes,aes(xmin=xmin, xmax=xmax, ymin=ymin,ymax=ymax), 
            inherit.aes = FALSE, 
            color='red', 
            fill=NA,
            size=1.5
  )

```

